---
title: "R Notebook"
output: html_notebook
---
```{r}
library(SBpitch)
library(StatsBombR)
library(tidyverse)
library(ggsoccer)
library(gt)
library(patchwork)
```
load libraries

```{r}
devtools::install_github("statsbomb/StatsBombR")
```
Download statsbomb package from github

```{r}
free_comps <- FreeCompetitions()
Comp <- FreeCompetitions() %>%
  filter(competition_name == "FIFA World Cup") 
Matches <- FreeMatches(Comp) 
StatsBombData <- StatsBombFreeEvents(MatchesDF = Matches, Parallel = FALSE)
StatsBombData <-  allclean(StatsBombData) 
StatsBombData <- goalkeeperinfo(StatsBombData)
StatsBombData <- shotinfo(StatsBombData)
StatsBombData <- formatelapsedtime(StatsBombData)
mensWC <- StatsBombData
```
Read in data from GitHub

```{r}
mensWC_filtered <- StatsBombData %>%
  select(period, minute, second, duration, under_pressure, counterpress, type.name, play_pattern.name, player.name, position.name, shot.statsbomb_xg, shot.first_time, shot.aerial_won, shot.follows_dribble, shot.outcome.name, shot.type.name, goalkeeper.position.name, goalkeeper.type.name, goalkeeper.outcome.name, foul_committed.advantage, foul_committed.penalty, foul_committed.type.name, foul_won.advantage, foul_won.defensive, foul_won.penalty, location.x, location.y, shot.open_goal, shot.one_on_one, DistToGoal, DistToKeeper, AngleToGoal, AngleToKeeper, distance.ToD1, distance.ToD2, player.name.GK, location.x.GK, location.y.GK, AttackersBehindBall, DefendersBehindBall, DefendersInCone, InCone.GK)
```
Select variables for analysis

```{r}
StatsBombCloseShots <- StatsBombFiltered %>%
  filter(type.name == "Shot", shot.type.name != "Penalty") %>%
  filter((location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62))) 

create_Pitch() +
  geom_point(data = StatsBombCloseShots, aes(x = location.x, y = location.y, color = shot.statsbomb_xg)) +
scale_color_gradient(low = "darkorange", high = "darkblue")
```
Expected Goals for shots within the 18


```{r}
StatsBombPenalties <- StatsBombFiltered %>%
  filter(type.name == "Foul Won") 
StatsBombPenalties %>%
  filter((location.x < 20) | (location.x > 100)) %>%
  ggplot(aes(x = location.x, y = location.y)) +
  geom_point() +
  theme_bw()
```

```{r}
devtools::install_github("FCrSTATS/SBpitch")
```
download SBpitch data

```{r}

closePenalties <- StatsBombPenalties %>%
  filter((location.x < 22) | (location.x > 98), ((location.y >= 14) & (location.y <= 66))) 

create_Pitch() +
  geom_point(data = closePenalties, aes(x = location.x, y = location.y)) +
  theme_bw() +
  labs(title = "Fouls Won Within and Around the 18",
       subtitle = "Men's World Cup 2018",
       x = " ",
       y = " ",
       caption = "Data from Statsbomb")
  
```
Fouls Won close to and within the 18

```{r}
StatsBombShots <- StatsBombFiltered %>%
  filter(type.name == "Shot")

ggplot()+
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point(data = StatsBombShots, aes(x = location.x, y = location.y, color = shot.statsbomb_xg, shape = (shot.outcome.name == "Goal"))) +
  scale_color_gradient(low = "darkorange", high = "darkblue")

ggplot()+
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point(data = StatsBombCloseShots, aes(x = location.x, y = location.y, color = shot.statsbomb_xg)) +
  scale_color_gradient(low = "darkorange", high = "darkblue", name = "Statsbomb Expected Goals") +
  labs(title = "Expected Goals for Open Play Shots",
       subtitle = "Men's World Cup 2018",
       caption = "Data Source: Statsbomb") 
  
  
```
Shots taken in mens world cup 2018

```{r}
Comp2 <- FreeCompetitions() %>%
  filter((competition_name == "FIFA World Cup") | (competition_name == "Champions League") | (competition_name == "La Liga") | (competition_name == "Premier League"))
Matches2 <- FreeMatches(Comp2) 
StatsBombData2 <- StatsBombFreeEvents(MatchesDF = Matches2, Parallel = FALSE)
StatsBombData2 <- allclean(StatsBombData2) 
StatsBombData2 <- goalkeeperinfo(StatsBombData2)
StatsBombData2 <- shotinfo(StatsBombData2)
StatsBombData2 <- formatelapsedtime(StatsBombData2)
allMens <- StatsBombData2
```
Read in data for all of the mens data availible 

```{r}
Comp3 <- FreeCompetitions() %>%
  filter(competition_name == "Women's World Cup")
Matches3 <- FreeMatches(Comp3) 
WomensWC <- StatsBombFreeEvents(MatchesDF = Matches3, Parallel = FALSE)
WomensWC <- goalkeeperinfo(WomensWC)
WomensWC <- allclean(WomensWC)
WomensWC <- shotinfo(WomensWC)
```
Download Women's World Cup Data

```{r}
allMens_filtered <- allMens %>%
  select(period, minute, second, duration, under_pressure, counterpress, type.name, play_pattern.name, player.name, position.name, shot.statsbomb_xg, shot.first_time, shot.aerial_won, shot.follows_dribble, shot.outcome.name, shot.type.name, goalkeeper.position.name, goalkeeper.type.name, goalkeeper.outcome.name, foul_committed.advantage, foul_committed.penalty, foul_committed.type.name, foul_won.advantage, foul_won.defensive, foul_won.penalty, location.x, location.y, shot.open_goal, shot.one_on_one, DistToGoal, DistToKeeper, AngleToGoal, AngleToKeeper, distance.ToD1, distance.ToD2, player.name.GK, location.x.GK, location.y.GK, AttackersBehindBall, DefendersBehindBall, DefendersInCone, InCone.GK)

womensWC_filtered <- WomensWC %>%
  select(period, minute, second, duration, under_pressure, counterpress, type.name, play_pattern.name, player.name, position.name, shot.statsbomb_xg, shot.first_time, shot.aerial_won, shot.follows_dribble, shot.outcome.name, shot.type.name, goalkeeper.position.name, goalkeeper.type.name, goalkeeper.outcome.name, foul_committed.advantage, foul_committed.penalty, foul_committed.type.name, foul_won.advantage, foul_won.defensive, foul_won.penalty, location.x, location.y, shot.open_goal, shot.one_on_one, DistToGoal, DistToKeeper, AngleToGoal, AngleToKeeper, distance.ToD1, distance.ToD2, player.name.GK, location.x.GK, location.y.GK, AttackersBehindBall, DefendersBehindBall, DefendersInCone, InCone.GK)
```
Select variables needed for analysis from both new datasets

```{r}
womensWC_filtered %>%
  filter(shot.type.name == "Penalty", period != 5) %>%
  nrow()

womensWC_filtered %>%
  filter(shot.type.name == "Penalty", period != 5, shot.outcome.name == "Goal") %>%
  nrow()

mensWC_filtered %>%
  filter(shot.type.name == "Penalty", period != 5) %>%
  nrow()

mensWC_filtered %>%
  filter(shot.type.name == "Penalty", shot.outcome.name == "Goal", period != 5) %>%
  nrow()

allMens_filtered %>%
  filter(shot.type.name == "Penalty", period != 5) %>%
  nrow()

allMens_filtered %>%
  filter(shot.type.name == "Penalty", shot.outcome.name == "Goal", period != 5) %>%
  nrow()
```

```{r}
penaltyKicksWon <- allMens_filtered %>%
  filter(minute <= 120, foul_committed.penalty == TRUE)

create_Pitch() +
  geom_point(data = penaltyKicksWon, aes(x = location.x, y = location.y))
```

```{r}
mensWC_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.type.name != "Penalty") %>%
  nrow()

mensWC_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.outcome.name == "Goal", shot.type.name != "Penalty") %>%
  nrow()


womensWC_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.type.name != "Penalty") %>%
  nrow()

womensWC_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.outcome.name == "Goal", shot.type.name != "Penalty") %>%
  nrow()


allMens_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.type.name != "Penalty") %>%
  nrow()

allMens_filtered %>%
  filter(type.name == "Shot", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.outcome.name == "Goal", shot.type.name != "Penalty") %>%
  nrow()
```

```{r}
boxShotsMWC <- mensWC_filtered %>%
  filter(type.name == "Shot", shot.type.name != "Penalty", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.statsbomb_xg >= .5)
  
above0.5MWC <- ggplot()+
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point(data = boxShotsMWC, aes(x = location.x, y = location.y)) +
  labs(title = "Men's World Cup")



boxShotsAllMens <- allMens_filtered %>%
  filter(type.name == "Shot", shot.type.name != "Penalty", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.statsbomb_xg >= .5)
  
above0.5allMens <- ggplot()+
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point(data = boxShotsAllMens, aes(x = location.x, y = location.y)) +
  labs(title = "All Men's Play")



boxShotsWWC <- womensWC_filtered %>%
  filter(type.name == "Shot", shot.type.name != "Penalty", (location.x < 18) | (location.x > 102), ((location.y >= 18) & (location.y <= 62)), shot.statsbomb_xg >= .5)
  
above0.5WWC <-ggplot()+
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point(data = boxShotsWWC, aes(x = location.x, y = location.y)) +
  labs(title = "Women's World Cup")

above0.5WWC + above0.5MWC 
```

```{r}
write.csv(mensWC_filtered,'mensWC.csv')
```


```{r}
penaltyShootoutMensWC <- mensWC_filtered %>%
  filter(period == 5)

penaltyShootoutWomensWC <- womensWC_filtered %>%
  filter(period == 5)

penaltyShootoutAllMens <- allMens_filtered %>%
  filter(period == 5)
```

```{r}
openShotsAllMens <- allMens_filtered %>%
  filter(type.name == "Shot", DefendersBehindBall < 4, AngleToGoal >= 70 & AngleToGoal <= 110, shot.type.name == "Penalty", period != 5)

openShotsAllMens %>%
  filter(shot.outcome.name == "Goal") %>%
  nrow()

openShotsAllMens %>%
  ggplot(aes(x = DistToGoal)) +
  geom_histogram() +
  theme_bw()

openShotsAllMens %>%
  ggplot(aes(x = location.x, y = location.y)) +
  annotate_pitch(dimensions = pitch_statsbomb)+
  theme_pitch()+
  coord_flip(xlim = c(55, 120),
             ylim = c(-12, 105)) +
  geom_point() 
```


```{r}
mensWCRegPKShots <- mensWC_filtered %>%
  filter(type.name == "Shot", shot.type.name == "Penalty", period != 5) 

mensWCRegPKFouls <- mensWC_filtered %>%
  filter(period != 5, foul_won.penalty == TRUE) 
```


